% % % % % % % % % % % % % % % % % % % %
%                                     %
% Towards Exascale Molecular Dynamics %
%                                     %
% % % % % % % % % % % % % % % % % % % %
%          Padraig O Conbhui          %
% % % % % % % % % % % % % % % % % % % %

\documentclass[12pt,a4paper]{report}

\input{macros}
\input{packages}
\input{styles}

\begin{document}


%
% Begin roman numerals
%

\pagenumbering{roman}


%
% Make front page
%

\makeEPCCtitle

\thispagestyle{empty}

\vspace{12cm}

\begin{center}

\large{MSc in High Performance Computing}

\large{The University of Edinburgh}

\large{Year of Presentation: \the\year}

\end{center}

\newpage

%
% Empty page
%
\thispagestyle{empty}
\clearpage\mbox{}\clearpage


%
% Project Abstract
%

\begin{abstract}
Computer experiments are becoming increasingly important in
scientific research.
%
In fields such as material science and biophysics, computer models
of atomic and molecular systems present a rich source of
information.
%
Molecular Dynamics (MD) is one tool for generating these computer models.

There is a constant strive to simulate ever larger molecular systems
using MD.
%
Given a simple formulation of an atomic system suggests a
$\bigO{N^2}$ algorithm, increasing simulation sizes is not
a straightforward task.
%
To this end, many approximations of they physics involved exist that
can dramatically reduce the time complexity of these simulations.
%
However, these approximations represent potential sources of error
in the already chaotic equations of motion for these systems.

This work attempts to determine just how far parallelisation alone
can improve an MD application.
%
In particular, it strives to find where the bottlenecks exist for
four parallel schemes which integrate the equations of motion for
an MD simulation without approximations.
%
Two of these schemes are the familiar \replicateddata{} and
\systolicloop{} schemes.
%
The third is referred to in this work as the \sharedandreplicateddata{}
scheme, representing a hybrid replicated and shared memory scheme.
%
The fourth, presented here, is referred to as the \replicatedsystolicloop{},
which is based on the \systolicloop{} scheme, taking inspiration from
the \replicateddata{} scheme by using replica systolic loops to
reduce the number of pulses any one systolic loop performs.

The bottleneck in the \replicateddata{} scheme was found to be
the amount of data being transferred and the number of MPI processes
being communicated to.
%
This was addressed by the \sharedandreplicateddata{} scheme by
using shared memory to reduce the number of MPI communications
necesary.
%
This scheme also addressed some of the memory usage issues involved
with the \replicateddata{} scheme.

The bottleneck in the \systolicloop{} scheme was found to be the
number of pulses performed, as a result of communication latency.
%
This was addressed by the \replicatedsystolicloop{} scheme by
using replica loops to reduce the number of systolic pulses each
loop performed.

\end{abstract}


% Restart numbering
\pagenumbering{roman}


%
% Content list pages
%

\tableofcontents
\listoftables
\listoffigures


%
% Title page
%

\begin{titlepage}
\vspace*{2in}
% an acknowledgements section is completely optional but if you decide
% not to include it you should still include an empty {titlepage}
% environment as this initialises things like section and page numbering.

%
% Acknowledgements
%

\section*{Acknowledgements}

I would like to thank my supervisor, Dr. Antonia Collis, without whom
I would still be completely incapable of writing a reasonable scientific
document.
%
Her guidance throughout the project, steering me back on track when I
needed steering, and giving me the advice I needed when I needed it,
were instrumental to the existance of this document.

Thanks go out to friends and family for their love and support
over the last 12 months. 

I must also thank my parents, in particular,
for bankrolling me all the way here.



\end{titlepage}


% Start regular page numbering
\pagenumbering{arabic}


%
% Include chapters
%

\include{introduction/introduction}
\include{background/background}
\include{methodology/methodology}
\include{parallel_implementation/parallel_implementation}
\include{conclusion/conclusion}

\bibliographystyle{unsrt}
\bibliography{bibliography}

\end{document}
